Literature Review

### 2018-07-10 LiteratureReview/LiteratureReview3000.tex
Updated Literature Review - Introduction Done

Alzheimer's Disease (AD) is a neurodegenerative disease in which the hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits. In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties. \newline

@@

According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of Dementia in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population. This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis. They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars. \newline

@@

Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. \newline

@@

According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word-find difficulty. They may substitute general terms for more specific terms. They may avoid the use of specific names of acquaintances. There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc. Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language. They will often use general-use phrases such as "that thing" and "you know what I mean," and prefers general pronouns rather than names. With severe impairment, sufferers may not even recall names of closer friends and family. Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur. Stereotypy of speech occurs; echolalia and automatic speech typically precede mutism.\newline

@@

% In the paragraph below, cite Goodglass and Kaplan (1983) and/or Goodglass (2000)
In assessing whether a client has AD, part of a clinicans job is to assess langugage capability. As described above, AD has a significant impact on language and a clinician has access to tools designed to test lanaguage capabilities. An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{} . In this task participants are asked to describe a picture presented to them in as much detail as possible. The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood. It was originally designed to assess Aphasia, but has shown itself to be useful in the diagnosis of AD as well \cite{GilesPattersonHodges1995}\newline

@@

The study of language decline in Dementia is not a new field of study. The focus of this work has been in understanding in which areas of language are impacted by dementia. Emery \cite{Emery2000} completed a literature review in this area and structured her review into four levels of language. There are four levels of language according to this perspective: Phonology, Morphology, Syntax and Semantics and her review looked at the idea that language and the processes involved in language are hierarchical. She proposes that language goes from simple units of construction, and build layers of complexity and sophistication. She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics. Emery states that you can see language decline as Hierarchical. According to a semiotic model there are four ranks of language based on their cognitive complexity. These ranks are Phonology, Morphology, Syntax and Semantics. She conducted a review studies which looked at language decline from the perspective of these hierarchical ranks. Her conclusion is that language decline is related to the complexity of the language task given to a participant and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.\newline

@@

There are a number of classical studies which look at language decline in those with dementia. One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan. They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease. Another \newline

@@

A lot of research since Emery's literature review \cite{Emery2000} have looked supported her ideas. \newline

@@

It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD. However, one of the costs of analysing in this way is the idea that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders, to facilitate the process of collecting data and analysis. The field of machine learning and natural language processing could be used to improve the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline.  
The purpose of this review is to seek to understand what developments have been made in the use of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and Alzheimer's Disease (AD). A search of the literature was conducted mainly from a psychological perspective using ProQuest (PsychArticles), SCOPUS, Web of Science. The following results were found (Table 1). All papers were the reviewed for relevance by reading the abstract and full text where appropriate. An additonal search through references of shortlisted papers was also conducted. Papers were included where there was an attempt to classify participants as MCI, AD or Healthy using speech and/or language by using Machine Learning techniques. We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English. This resulted in 21 journal articles and/or conference papers which form this review.

@@

\section {Types of Language Assessment}
\subsection{Picture Description Tasks}
One of the most commonly used tasks to measure lexical analysis is the Picture Description task. In dementia research, the most common task picture description task is the Cookie Theft picture. This does very well in giving a valid dataset of language but suffers somewhat from ecological validity. In essence, when completing this task, the participant is generating speech to describe the picture but does not use some of the skills generally used in every day interactions. If the goal is to a way 

>>>>

Alzheimer's Disease (AD) is a neurodegenerative disease in which the brain develops neurofibrillary tangles and neuritic plaques along with the loss of cortical neurons and synapses. The hallmark clinical symptoms are cognitive deficits such as problems with episodic and semantic memory, organising and planning, difficulties with language and visuospatial deficits\cite{AmericanPsychiatricAssociation2013}. In addition, these symptoms are often accompanied by emotional difficulties such as depression and irritability and behavioural difficulties. Whilst there are several variants of Dementia, AD remains the most common type of Dementia and will be the focus of this review. \newline

@@

According to a recent report commissioned by the Alzheimer's Society in 2015, they estimate the prevalence of AD in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over and 1 in 79 in the general population. They estimate an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis and that the total impact of AD (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. A recent report suggests that in 2015 there were 46 million people with a diagnosis of AD and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of AD in 2018 is estimated to be in the region of one trillon US dollars. \newline

@@

Current thinking suggests that the cognitive deficits associated with AD often begin before the clinical symptoms of the disease become apparent. Researchers propose that neurofibrillary tangles and other associated physiological effects of AD develop over time until a threshold is reached and clinical symptoms become apparent \cite{Nestor2006}. If this is the case, it should be possible to detect subtle cognitive changes in language and memory before a clinical diagnosis can be formed. Given that language is less intrusive to test and requires a lot of the cognitive processes that may be impacted by AD, a lot of research has focused on the decline in the use of language in those with AD.  \newline

@@

One of the most famous pieces of research was by Berisha and Liss (2015) \cite{Berisha2015} who examined speeches and public interviews of former US president Ronald Reagan. They found that Reagan's speeches towards the end of his presidency suffered from difficulties in word-finding, inappropriate phrases and uncorrected sentences which are hallmarks of language deterioration associated with Alzheimer's Disease. Another classical study by Snowdon et al (1996) \cite{Snowdon1996} looked at whether linguistic ability in early life was asoociated with cognitive function and AD in later life. They found that idea density (defined as the number of expressed propositions divided by the number of words) was a key predictor in predicting whether nuns would go on to develop AD in later life. They found that those who would go on to develop AD all had low idea density in early life and they found no AD present in those with high idea density. \newline

@@

The range of language deficits in those who suffer with AD are wide and varied and differ as the disease develops. According to the DSM 5 \cite{AmericanPsychiatricAssociation2013}, those with Mild dementia suffer from noticeable word finding difficulty. They may substitute general terms for more specific terms and may avoid the use of specific names of acquaintances. There may be grammatical errors involving subtle omission or incorrect use of articles, prepositions, auxiliary verbs, etc. Those who have progressed from Mild to Major depression also have difficulties with expressive or receptive language. They will often use general-use phrases such as "that thing" and "you know what I mean" and prefers general pronouns rather than names. With severe impairment, sufferers may not even recall names of closer friends and family. Idiosyncratic word usage, grammatical errors, and spontaneity of output and economy of utterances occur. Echolalia (meaningless repetition of another person's spoken words) and automatic speech typically precede mutism. With the wide range of deficits someone with AD can suffer, it makes sense to try to categorise these deficits in some way.\newline

@@

Emery \cite{Emery2000} completed a literature review in this area and she looked at four levels of language: Phonology, Morphology, Syntax and Semantics. Her review looked at the idea that language and the processes involved in language are hierarchical. She proposes that language goes from simple units of construction, and build layers of complexity and sophistication. She found that people with AD generally had intact Phonology and Morphology but more impaired Syntax and Semantics. Emery concludes that language decline is hierarchical and is related to the complexity of the language task given to a participant. She and that language decline is hierarchical in that the language forms we learn last (the most complex language forms) are the first to deteriorate.\newline

@@

It is clear from both the clinical diagnostic criteria and supporting research that language is impacted in those with AD. However, one of the costs of analysing language is that there is a huge burden on trained practitioners, be it clinical psychologists, audio transcribers and text encoders to facilitate the process of collecting data and analysis. The field of machine learning and natural language processing has been suggested as a way toimprove the accuracy and lessen the human cost of this research as well as provide new insights into the difficulties that AD suffer in terms of language decline. \newline

@@

The purpose of this review is to seek to understand what developments have been made in the use of Machine Learning and Natural Language processing to aid the diagnosis of Mild Cognitive Impairment (MCI) and AD. A search of the literature was conducted using ProQuest (PsychArticles), SCOPUS, Web of Science. The following results were found (Table 1). All papers were then reviewed for relevance by reading the abstract and full text where appropriate and a shortlist was compiled. An additonal search through references of shortlisted papers was also conducted. Papers were included where researchers used machine learning to classify participants as MCI, AD or Healthy using language. We excluded any papers that focused on other forms of dementia or cognitive impairment, as well as any papers in which the language being analysed was not English. This resulted in 21 journal articles and/or conference papers which form this review.

\section {Types of Language Assessment}
One of the key debates when looking at how to analyse language is the type of task provided to elicit language. In the literature  researchers have primarily focused on Picture Description tasks but have also suggested other ways in which we might collect data. \newline
\par
\subsection{Picture Description Tasks}
One of the most commonly used tasks to measure lexical analysis is the Picture Description task. In dementia research, the most common task picture description task is the Cookie Theft picture. This does very well in giving a valid dataset of language but suffers somewhat from ecological validity. In essence, when completing this task, the participant is generating speech to describe the picture but does not use some of the skills generally used in every day interactions. If the goal is to a way

In assessing whether a client has AD, part of a clinicans job is to assess langugage capability. As described above, AD has a significant impact on language and a clinician has access to tools designed to test lanaguage capabilities. An example of this is part of the Boston Diagnostic Aphasia Examination (BDAE), called the Boston Cookie Theft picture description task \cite{Kaplan2010} . In this task participants are asked to describe a picture presented to them in as much detail as possible. The picture itself depicts a familiar domestic scene and would not require participants to use any vocabulary beyond that learned in childhood. It was originally designed to assess Aphasia, but has shown itself to be useful in the diagnosis of AD as well \cite{Giles1996}\newline

@@

@@
